---
title: "Interactivity: Quarto & Observable JS"

format: 
  html:
    df-print: paged
    toc: true
    echo: false
    warning: false
    messages: false
    code-summary: "Show the code"
    code-tools: true
---

```{r}
library(tidyverse)
library(haven)
foo_df <- read_sav("data/student_satisfaction_test-data_from_qualtrics.sav")
satisfaction_df <- foo_df |> janitor::clean_names()
satisfaction_df <- read_csv(here::here("data", "student_satisfaction_test-data_from_qualtrics.sav"), 
    col_names = FALSE, skip = 3)
# my_df_colnames_and_questions <- 
#   read_csv("data/qualtrics_student_satisfaction-test-data.csv", 
#            n_max = 1) |> 
#   janitor::clean_names()
# my_colnames_vec <- colnames(my_df_colnames_and_questions)
# colnames(satisfaction_df) <- my_colnames_vec
trade_df <- read_csv("data/data.csv")
trade_df <- trade_df |> 
  filter(str_detect(business, "(Motor|beverage|Electronics)")) |> 
  mutate(sales = as.integer(sales))
```

```{r}
foo_df <- foo_df |> 
  janitor::clean_names() |> 
  mutate(across(is.labelled, ~ as_factor(.x), .names = "{.col}_label")) |>
  # select(sort(tidyselect::peek_vars())) |> 
  select(response_id, ends_with("_label"), -c(status_label, finished_label)) |> 
  rename(q15 = q15_label) |> 
  mutate(q15 = as.double(q15)) |>
  pivot_longer(-c(response_id, q15), names_to = "my_question", values_to = "response") |> 
  pull(response) |> 
  levels()
foo_df
```

```{r}
#| eval: false
my_q1q4_levels <- rev(c("Extremely helpful", 
                   "Moderately helpful", 
                   "Slightly helpful", 
                   "Neither helpful nor unhelpful",
                   "Slightly unhelpful", 
                   "Moderately unhelpful", 
                   "Extremely unhelpful" ))

my_df_tall <- satisfaction_df |> 
  select(response_id, q1:last_col()) |> 
  pivot_longer(-c(response_id, q15), names_to = "my_question", values_to = "response") |> 
  filter(my_question %in% c("q1", "q4"))  |> 
  mutate(response = factor(response, levels = my_q1q4_levels)) |> 
  mutate(rnumber = as.numeric(response))
```

```{r}
ojs_define(trade_df_ojs = trade_df)
# ojs_define(my_df_tall_ojs = my_df_tall)
ojs_define(my_df_tall_ojs = foo_df)
```

## Interactivity (Observablejs)

This is a very basic example of diplaying an interactive visualization using ObservableJS.

```{ojs}
//| panel: input
//| layout-ncol: 1

// viewof my_question = Inputs.radio(
//  ["q1", "q4"], 
//  {  value: "q1", 
//     label: html`<b>Choose Question</b>`,
//     multiple: true
// })

viewof radio_2 = 
Inputs.radio(
  new  Map([
    ["Question One",  'q1'], 
    ["Question Four", 'q4']
  ]),
  {  label: html`<b>Survey results summarized:</b>`,
     value: 'q1',
     multiple: true
})
```

::: panel-tabset
## Plot

```{ojs}
//| code-fold: true
//| echo: fenced

my_survey_filtered = my_survey
  .filter(aq.escape(d => d.my_question == radio_2))
  .groupby('my_question', 'response', 'rnumber')
  .count()
  .orderby('my_question', aq.desc('rnumber'))
  
Plot.plot({
  marginRight: 140,
  y: {
    axis: null,
    domain: d3.sort(my_survey_filtered, d => -d.rnumber).map(d => d.response)
  },
  x: {
    grid: true
  },
 marks: [
    Plot.barX(my_survey_filtered, {y: "response", x: "count"}),
    Plot.text(my_survey_filtered, {
      x: "count",
      y: "response",
      text: "response",
      textAnchor: "start",
      dx: 6
    })
  ]
})
```

## Data

```{ojs}
my_survey_filtered.view()
```
:::

## Download

```{ojs}
downloadButton(my_survey.objects(), "Qualtrics_test_data.csv")
```

## Transpose to arquero

Behind the scenes, when rendering this report, I use `{r}` code-chunks to import and wrangle the data. Just like I did in [*code how you like*](basic.html), I'm importing and wrangling my data with a quarto-friendly code language. Then I *transpose* my data frame into an OJS object.

```{ojs}
//| echo: fenced
my_data = aq.from(transpose(trade_df_ojs))
my_survey = aq.from(transpose(my_df_tall_ojs))
```

::: callout-note
## Reveal code: click the `</> Code` {{< fa solid code >}} icon

This page reveals the observablejs `{ojs}` code-chunks. However, this page demonstrates using `{r}` and `{ojs}` together in the same quarto-document. You can scroll to the top and click the `</> Code` {{< fa solid code >}} icon to see all the code. Or click the GitHub icon {{< fa brands github >}} in the navigation bar. Quarto makes it easy to include these features.
:::

## One more example

The next code-chunk shows a very simple example of ObservableJS `{Plot}` code for visualization. This example *is not interactive,* to show the remarkable similarity to `{ggplot2}` syntax.

plot

```{ojs}
//| echo: fenced
Plot.plot({
  marks: [
    Plot.ruleY([0]),
    Plot.lineY(my_data_wrangled, {
      x: "date",
      y: "sales",
      stroke: "business"
    }),
  ]
})
```

::: callout-important
## ↥ Look up!

Notice the crispness and clarity of the simple non-interactive plot above. [Observable Plot](https://observablehq.com/@observablehq/plot) shows well on the web!
:::

## Arquero data wrangling

::: gray
**Derive date**

Aside from the `{dplyr}` wrangling that I did in the background of this code notebook, I still needed to parse the date fields. Arquero to the rescue.

**Arquero**! Now what is that? *Observable JavaScript* is its own coding language and we can minimize the learning curve by keeping our data wrangling within our favorite coding language. **Then** we can focus on using the [{Plot}](https://observablehq.com/plot) grammar and syntax, which is quite similar to `{ggplot2}` -- see the above example. However, [Observable JS](https://quarto.org/docs/interactive/ojs/) is a complete and rich coding language of its own. But just like [{`shiny`}](https://quarto.org/docs/interactive/shiny/), we want to learn just enough to get the job done. Still, there's always an edge case where a little data wrangling is optimal. This is where [`{arquero}`](https://observablehq.com/@uwdata/introducing-arquero) comes in. Arquero is used for data transformation and is easy to learn because it was inspired by `{dplyr}`. *Show the code* for the first plot on this page and you'll see the similarity to `{dplyr}`.

Below I parse a data type using `{arquero}`.
:::

```{ojs}
//| echo: fenced
my_data_wrangled = my_data
  .derive({  date: d =>  op.parse_date(d.date) })
```

## Observablejs libraries

By convention, observable libraries are loaded at the bottom of a code notebook. You can read more about observablejs at Observablehq.com. By way of a teaser, I didn't have to use `{r}` to wrangle the retail data. I could have *imported the data directly from observable*. Such as ....

\> `import { data, advancedDate } from "@observablehq/monthly-retail-trade-report-data"`

But on this page, I'm demonstrating how we leveraged `{R}` and `{ojs}` code with Quarto, so that I can present useful interactivity and demonstrate multi-lingual coding using a Quarto notebook.

```{ojs}
//| echo: fenced
import { aq, op } from '@uwdata/arquero'
// import { data, advancedDate } from "@observablehq/monthly-retail-trade-report-data"
import {button as downloadButton} from "@jeremiak/download-data-button"
```
