---
title: "Interactivity: Quarto & Observable JS"

format: 
  html:
    df-print: paged
    toc: true
    echo: false
    warning: false
    messages: false
    code-summary: "Show the code"
    code-tools: true
---

```{r}
library(tidyverse)
library(haven)
satisfaction_df <- read_sav("data/student_satisfaction_test-data_from_qualtrics.sav") |> 
  janitor::clean_names() |> 
  mutate(across(is.labelled, ~ as_factor(.x), .names = "{.col}_label")) |> 
  select(response_id, q1_label, q4_label, q5_label, q8_label, q10_label, q12_label) |> 
  pivot_longer(-response_id, names_to = "my_question", values_to = "response") |> 
  mutate(rnumber = as.numeric(response))

trade_df <- read_csv("data/data.csv")
trade_df <- trade_df |> 
  filter(str_detect(business, "(Motor|beverage|Electronics)")) |> 
  mutate(sales = as.integer(sales))
```

```{r}
ojs_define(trade_df_ojs = trade_df)
ojs_define(my_df_tall_ojs = satisfaction_df)
```

This is a very basic example of displaying an interactive visualization using ObservableJS. Select a Question, then repeat after choosing between the *Plot* or *Data* tabs. You'll see the visualizations are reactive to your selections.

```{ojs}
//| panel: input
//| layout-ncol: 1

viewof radio_2 = 
Inputs.radio(
  new  Map([
    ["One",    'q1_label'], 
    ["Four",   'q4_label'],
    ["Five",   'q5_label'],
    ["Eight",  'q8_label'],
    ["Ten",    'q10_label'],
    ["Twelve", 'q12_label']
  ]),
  {  label: html`<b>Question:</b>`,
     value: 'q1_label',
     multiple: true
})
```

::: panel-tabset
## Plot

```{ojs}
//| code-fold: true
//| echo: fenced

my_survey_filtered = my_survey
  .filter(aq.escape(d => d.my_question == radio_2))
  .groupby('my_question', 'response', 'rnumber')
  .count()
  .orderby('my_question', 'rnumber')
  
Plot.plot({
  marginRight: 140,
  y: {
    axis: null,
    domain: d3.sort(my_survey_filtered, d => -d.rnumber).map(d => d.response).reverse()
  },
  x: {
    grid: true
  },
 marks: [
    Plot.barX(my_survey_filtered, {y: "response", x: "count"}),
    Plot.text(my_survey_filtered, {
      x: "count",
      y: "response",
      text: "response",
      textAnchor: "start",
      dx: 6
    })
  ]
})
```

## Data

```{ojs}
my_survey_filtered.view()
```
:::

::: {.callout-tip collapse="true"}
## Workflow: How'd we do that?

The above example uses R and the {haven} package to import labeled Qualtrics data. An SPSS *sav* data file with raw data, variable and value labels is uploaded via `read_sav`, warngled with {dplyr}, and then visuazlized with client-side interactive ObservableJS. All made possible by R and Quarto.

All the code is in the code repo, or see the \</\> icon at the top of this page, then View source.
:::

## Transpose to arquero

Behind the scenes, when rendering this report, I use `{r}` code-chunks to import and wrangle the data. Just like I did in [*code how you like*](basic.html), I'm importing and wrangling my data with a quarto-friendly code language. Then I *transpose* my data frame into an OJS object.

```{ojs}
//| echo: fenced
my_data = aq.from(transpose(trade_df_ojs))
my_survey = aq.from(transpose(my_df_tall_ojs))
```

::: callout-note
## Reveal code: click the `</> Code` {{< fa solid code >}} icon

This page reveals the observablejs `{ojs}` code-chunks. However, this page demonstrates using `{r}` and `{ojs}` together in the same quarto-document. You can scroll to the top and click the `</> Code` {{< fa solid code >}} icon to see all the code. Or click the GitHub icon {{< fa brands github >}} in the navigation bar. Quarto makes it easy to include these features.
:::

## One more example

The next code-chunk shows a very simple example of ObservableJS `{Plot}` code for visualization. This example *is not interactive* but it demonstrates the remarkable similarity to `{ggplot2}` syntax.

plot

```{ojs}
//| echo: fenced
Plot.plot({
  marks: [
    Plot.ruleY([0]),
    Plot.lineY(my_data_wrangled, {
      x: "date",
      y: "sales",
      stroke: "business"
    }),
  ]
})
```

::: callout-important
## ↥ Look up!

Notice the crispness and clarity of the simple non-interactive plot above. [Observable Plot](https://observablehq.com/@observablehq/plot) shows well on the web!
:::

## Arquero data wrangling

::: gray
**Derive date**

Aside from the `{dplyr}` wrangling that I did in the background of this code notebook, I still needed to parse the date fields. Arquero to the rescue.

**Arquero**! Now what is that? *Observable JavaScript* is its own coding language and we can minimize the learning curve by keeping our data wrangling within our favorite coding language. **Then** we can focus on using the [{Plot}](https://observablehq.com/plot) grammar and syntax, which is quite similar to `{ggplot2}` -- see the above example. However, [Observable JS](https://quarto.org/docs/interactive/ojs/) is a complete and rich coding language of its own. But just like [{`shiny`}](https://quarto.org/docs/interactive/shiny/), we want to learn just enough to get the job done. Still, there's always an edge case where a little data wrangling is optimal. This is where [`{arquero}`](https://observablehq.com/@uwdata/introducing-arquero) comes in. Arquero is used for data transformation and is easy to learn because it was inspired by `{dplyr}`. *Show the code* for the first plot on this page and you'll see the similarity to `{dplyr}`.

Below I parse a data type using `{arquero}`.
:::

```{ojs}
//| echo: fenced
my_data_wrangled = my_data
  .derive({  date: d =>  op.parse_date(d.date) })
```

## Observablejs libraries

By convention, observable libraries are loaded at the bottom of a code notebook. You can read more about observablejs at Observablehq.com. By way of a teaser, I didn't have to use `{r}` to wrangle the retail data. I could have *imported the data directly from observable*. Such as ....

\> `import { data, advancedDate } from "@observablehq/monthly-retail-trade-report-data"`

But on this page, I'm demonstrating how we leveraged `{R}` and `{ojs}` code with Quarto, so that I can present useful interactivity and demonstrate multi-lingual coding using a Quarto notebook.

```{ojs}
//| echo: fenced
import { aq, op } from '@uwdata/arquero'
// import { data, advancedDate } from "@observablehq/monthly-retail-trade-report-data"
import {button as downloadButton} from "@jeremiak/download-data-button"
```
