---
title: "Iteration with {purrr}"

date-modified: 'today'
date-format: long

license: CC BY-NC
bibliography: references.bib
---

::: {.callout-warning collapse="true"}
## Review the functions page!

If you have not already read the [functions page](functions.html), do so before reading about {purrr}. The purrr package is about applying functions repetitively. You should have a good idea about what a function is and how to write a custom function.
:::

From the [functions page](functions.html) we learned scant definitions of *functional programming* and *functions*. We also learned about the special conditions of programming with {dplyr}, which demands a working understanding of *environment* and *data* variables.[^1] Now we want to apply our functions, row-by-row, to a data frame. We'll use the purrr package.

[^1]: Links and footnotes on the [functions page](functions.html) will lead to more detailed information on those topics.

Remember that in functional programming we're iterating, or recursing, without using FOR loops. For example, in the [regression page](regression.html#example-of-iterative-modeling-with-nested-categories-of-data), we saw an example of *nesting data frames* by category (i.e. by *gender*). After nesting, we have two subsetted data frames that are embeded within a parent-data-frame. The embedded data frames are contained within a *list-column*.

### `nest()`

We can subset by gender, creating two new subset data frames that we'll put in a new column with the variable name: *data*. This subsetting is accomplished with the `nest()` function.

::: column-margin
<iframe width="300" height="200" src="https://www.youtube.com/embed/kZf11zbVpr0" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" allowfullscreen>

</iframe>
:::

```{r}
#| warning: false
#| message: false

library(moderndive)
library(broom)
library(tidyverse)

evals |> 
  janitor::clean_names() |> 
  nest(data = -gender)
```

## `map()`

Now, using {purrr} we can apply a function --- such as a linear model regression `lm()` --- over each row of the parent data frame. To do this, we use the `map_` class of functions. I say class of functions, because map allows us to define the returned data type. Sometimes we want to return a character data type (`map_char`), sometimes an integer (`map_int`), etc. In the case of a linear model, which is a list data-type, we'll use the generic `map()` function to apply an ***anonymous** function*.

::: callout-note
## Anonymous functions

An example of an anonymous function: `\(x) x + 1` or `function(x) x + 1`.
:::

::: column-margin
<iframe width="300" height="200" src="https://www.youtube.com/embed/QgasjZGhWlk" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" allowfullscreen>

</iframe>
:::

```{r}
#| warning: false
#| message: false

my_subset_df <- evals |> 
  janitor::clean_names() |> 
  nest(my_data = -gender) |> 
  mutate(my_fit_bty = map(my_data, function(my_data) 
                                lm(score ~ bty_avg, data = my_data)
                         )
        )
my_subset_df
```

### Look at the fitted object

When we look at the resulting `my_fit_bty` function, we see the kind of output we get from `lm()`

```{r}
my_subset_df$my_fit_bty
```

Earlier, we learn to use the `broom::tidy()` function to wrap a fitted model into a data frame[^2]. We can easily `tidy` just the first model.

[^2]: Having our regression model wrapped as a data frame means we use other {dplyr} functions to more easily manipulate our model output.

```{r}
tidy(my_subset_df$my_fit_bty[[1]])
```

### Named functions

To tidy all the models, we want to use the `map` function again. This time we use a named function, i.e. `tidy()`.

```{r}
my_subset_df |> 
  mutate(my_fit_bty_tidy = map(my_fit_bty, tidy))
```

The only difference between `my_fit_bty` and `my_fit_bty_tidy` is that `my_fit_bty_tidy` is a data frame. In fact, it's a nested just like `my_subset_df$data` which we nested in the first code-chunk of this page.

To unnest the data frames, use the `unnest()` function. I'll explain *named* and *anonymous* functions later.

```{r}
my_subset_df |> 
  mutate(my_fit_bty_tidy = map(my_fit_bty, tidy)) |> 
  unnest(my_fit_bty_tidy)
```

Now we have fitted model data, contained within a data frame that we can manipulate further. For example, let's just look at the p-values of `bty_avg`, by `gender`.

```{r}
my_subset_df |> 
  mutate(my_fit_bty_tidy = map(my_fit_bty, tidy)) |> 
  unnest(my_fit_bty_tidy) |> 
  select(gender, term, p.value) |> 
  filter(term == "bty_avg")
```

When we combine linear modeling with other broom functions such as `glance()` and `augment()`, then we can further our analysis and data manipulation.

## Anonymous functions

Anonymous functions have no name. Unlike a **named** function (e.g. `broom::tidy` above, or `make_scatterplot` from the [functions page](functions.html)). Below we **map** an anonymous function within `map()`. The first argument to map is a list or data frame. The next argument can be a named function or an anoymous function. Anonymous functions begin with the `tilde` and can use the .x argument to refer to the first argument. In the case below `.x` refers to the `data`.

.x refers to the first argument in the `map()` function. In this case, `.x` is a data frame, so we can use ggplot2 functions to generate our plot.

```{r}
my_subset_df_with_plots <- my_subset_df |> 
  mutate(my_plot = map(my_data, function(my_data) 
    my_data |> 
      ggplot(aes(bty_avg, score)) + 
      geom_point() +
      geom_smooth(method = lm, se = FALSE, formula = y ~ x)
                      )
         )
my_subset_df_with_plots
```

And now we can pull those plots

```{r}
#| eval: false
#| echo: true
my_subset_df_with_plots |> 
  pull(my_plot)
  
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-cap: 
#|   - "Female"
#|   - "Male"


my_subset_df_with_plots$my_plot[[1]]

my_subset_df_with_plots$my_plot[[2]]
```
